---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Summary

For a range of parameter values and all three alternative hypothesis, type I error rates are estimated via simulation. The estimates are based on 5,000 iterations and a sample size of 50. Where possible, exact tests are included for comparison. At this sample size, most tests are just above the .05 error rate.

```{r setUp, echo=FALSE, message=FALSE}
library(tidyverse, warn.conflicts= FALSE)
options(dplyr.summarise.inform = FALSE)

gaussian <- bind_rows(
  readRDS("results/gaussian_type_one.rds"),
  readRDS("results/gaussian_type_one_exact.rds")
)
gaussian <- gaussian %>% 
  filter(test %in% c("gaussian_mean_lr_test", "t.test")) %>%
  select(test, alt, param = mu, pvalue) %>%
  bind_rows(gaussian %>% 
              filter(test %in% c("gaussian_variance_lr_test", "varTest")) %>%
              select(test, alt, param = variance, pvalue))

gamma <- bind_rows(
  readRDS("results/gamma_type_one_rate.rds"),
  readRDS("results/gamma_type_one_scale.rds"),
  readRDS("results/gamma_type_one_shape.rds")
)
gamma <- gamma %>% 
  filter(test %in% c("gamma_rate_lr_test")) %>%
  select(test, alt, param = rate, pvalue) %>%
  bind_rows(gamma %>% 
              filter(test %in% c("gamma_scale_lr_test")) %>%
              select(test, alt, param = scale, pvalue),
            gamma %>% 
              filter(test %in% c("gamma_shape_lr_test")) %>%
              select(test, alt, param = shape, pvalue))

poisson <- readRDS("results/poisson_type_one.rds")
poisson <- poisson %>%
  select(test, alt, param = lambda, pvalue)


beta <- readRDS("results/beta_type_one.rds")
beta <- beta %>% 
  filter(test %in% c("beta_shape1_lr_test")) %>%
  select(test, alt, param = shape1, pvalue) %>%
  bind_rows(beta %>% 
              filter(test %in% c("beta_shape2_lr_test")) %>%
              select(test, alt, param = shape2, pvalue))

neg_binom <- bind_rows(
  readRDS("results/negative_binomial_type_one.rds"),
  readRDS("results/negative_binomial_type_one_exact.rds") %>%
    mutate(test = "negative_binomial_exact")
)
neg_binom <- neg_binom %>% 
  select(test, alt, param = p, pvalue)

geomet <- bind_rows(
  readRDS("results/geometric_type_one.rds"),
  readRDS("results/geometric_type_one_exact.rds") %>%
    mutate(test = "geometric_exact")
)
geomet <- geomet %>% 
  select(test, alt, param = p, pvalue)

expon <- readRDS("results/exponentail_type_one.rds")
expon <- expon %>% 
  select(test, alt, param = rate, pvalue)

typeI <- bind_rows(
  gaussian,
  gamma, 
  poisson, 
  beta,
  neg_binom,
  geomet,
  expon
) %>%
  drop_na()

mark_exact_likelihood <- function(test) {
  type <- case_when(
    str_detect(test, "t.test") ~ "exact",
    str_detect(test, "exact") ~ "exact",
    str_detect(test, "varTest") ~ "exact",
    str_detect(test, "_lr_test") ~ "approximate",
    TRUE ~ "ERROR")
  
  return(type)
}

typeI <- typeI %>%
  mutate(type = mark_exact_likelihood(test))
# typeI %>% distinct(type, test) %>% arrange(type, test) %>% print(N=Inf)

link_exact_to_likelihood <- function(test) {
  test2 <- case_when(
    str_detect(test, "t.test") ~ "gaussian_mean_lr_test",
    str_detect(test, "varTest") ~ "gaussian_variance_lr_test",
    str_detect(test, "geometric_exact") ~ "geometric_p_lr_test",
    str_detect(test, "negative_binomial_exact") ~ "negative_binomial_p_lr_test",
    str_detect(test, "_lr_test") ~ test,
    TRUE ~ "ERROR")
}

typeI <- typeI %>%
  mutate(test2 = link_exact_to_likelihood(test))
# typeI %>% distinct(type, test, test2) %>% arrange(type, test, test2) %>% print(N=Inf)

temp <- typeI %>%
  group_by(type, test2) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

ggplot(temp, aes(x = TypeI, y = test2, color = factor(type))) + 
  geom_point() +
  geom_vline(xintercept = .05) + 
  scale_x_continuous(breaks = seq(0, 1, .01), limits = c(0, .10)) + 
  labs(x = "Type I Error Rate", y = "Test", color = "Type")

rm(list = ls())


```
Exploring each test one by one, most tests are consistent across the entire parameter space and alternative hypotheses. The negative binomial has an area of poor type I error rate. The geometric tests did poorly overall.

# Successful Distributions
For a distribution, the likelihood ratio test works well if

* All tests achieve near .05 type I error rate over the entire parameter space.
* All tests achieve near .05 type I error across all alternative hypothesis.  

To support both points, two graphs are shown. Where possible, exact tests are included for comparison.

## Gaussian
```{r gaussainTypeI, echo=FALSE, message=FALSE, results='hide', warning=FALSE}

typeI <- bind_rows(
  readRDS("results/gaussian_type_one.rds") %>%
    mutate(type = "Likelihood"),
  readRDS("results/gaussian_type_one_exact.rds") %>%
    mutate(type = "Exact")
)

temp <- typeI %>%
  group_by(type, test, mu, variance) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = mu, y = factor(variance), fill = TypeI, label = TypeI, group = mu)) +
  geom_tile(height = .50, width = .75) +
  geom_text() +
  scale_x_continuous(breaks = seq(-3, 3, 1)) +
  scale_fill_gradient(low = "white", high = "red") +
  labs(title = "Overall Type I Error Rate", x = "mu", y = "variance", fill = "Type I Error") +
  facet_wrap(vars(test))

temp <- typeI %>%
  group_by(type, test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```


## Gamma
```{r gammaTypeI, echo=FALSE}

typeI <- bind_rows(
  readRDS(file = "results/gamma_type_one_rate.rds"),
  readRDS(file = "results/gamma_type_one_scale.rds"),
  readRDS(file = "results/gamma_type_one_shape.rds")
)

temp <- typeI %>%
  mutate(scale = round(scale, 2)) %>%
  group_by(test, alt, shape, rate, scale) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI))

temp_02 <- temp %>%
  mutate(TypeI = round(TypeI, 2)) %>%
  group_by(test, shape, rate) %>%
  summarise(TypeI = max(TypeI)) %>%
  ungroup() %>%
  arrange(desc(TypeI)) 

temp_02 %>%
  ggplot(aes(x = shape, y = factor(rate), fill = TypeI, label = TypeI, group = shape)) +
  geom_tile(height = .50, width = .50) +
  geom_text() +
  scale_x_continuous(breaks = seq(1, 10, 1)) + 
  scale_fill_gradient(low = "white", high = "red") +
  labs(x = "Shape", y = "Rate", fill = "Type I Error") + 
  facet_wrap(vars(test))

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```


## Poisson
```{r poissonTypeI, echo=FALSE}

typeI <- readRDS("results/poisson_type_one.rds")

temp <- typeI %>%
  group_by(test, lambda) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI))

ggplot(temp, aes(x = factor(lambda), y = TypeI)) +
  geom_point() +
  geom_hline(yintercept = .05) +
  scale_y_continuous(breaks = seq(0, 1, .01), limits = c(0, .10)) +
  labs(x = "Lambda", y = "Type I Error")

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```


## Beta
```{r betaTypeI, echo=FALSE}

typeI <- readRDS("results/beta_type_one.rds")

temp <- typeI %>%
  group_by(test, alt, shape1, shape2) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI))

temp_02 <- temp %>%
  mutate(TypeI = round(TypeI, 2)) %>%
  group_by(test, shape1, shape2) %>%
  summarise(TypeI = max(TypeI)) %>%
  ungroup() %>%
  arrange(desc(TypeI)) 

temp_02 %>%
  ggplot(aes(x = shape1, y = factor(shape2), fill = TypeI, label = TypeI, group = shape1)) +
  geom_tile(height = .50, width = .70) +
  geom_text() +
  scale_x_continuous(breaks = seq(1, 10, 1)) + 
  scale_y_discrete(breaks = seq(1, 10, 1)) + 
  scale_fill_gradient(low = "white", high = "red") +
  labs(x = "Shape", y = "Rate", fill = "Type I Error") +
  facet_wrap(vars(test))

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```


## Exponential
```{r exponentialTypeI, echo=FALSE}

typeI <- readRDS("results/exponentail_type_one.rds")

temp <- typeI %>%
  group_by(test, alt, rate) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI))

ggplot(temp, aes(x = factor(rate), y = TypeI)) +
  geom_point() +
  geom_hline(yintercept = .05) +
  scale_y_continuous(breaks = seq(0, 1, .01), limits = c(0, .10)) +
  labs(x = "P", y = "Type I Error")

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```

# Failures 
For a distribution, the likelihood ratio test is considered bad if

* Any test did not achieve near .05 type I error rate over any area parameter space.
* Any test did not achieve near .05 type I error across any alternative hypothesis.

It is possible for a test to "fail" and still be good approximation to the exact test.

## Negative Binomial
For most parameter values, the negative binomial test performs well. For values near 0 or 1, the test does not perform well for small number of target successes. The exact test does well where the likelihood test fails.

```{r negativeBonimialTypeI, echo=FALSE}

typeI <- bind_rows(
  readRDS("results/negative_binomial_type_one.rds") %>%
    mutate(type = "Likelihood"),
  readRDS("results/negative_binomial_type_one_exact.rds") %>%
    mutate(type = "Exact")
)

typeI <- typeI %>%
  filter(size <= 100)

temp <- typeI %>%
  group_by(type, test, alt, p, size) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

# summarize 3 alts into one row by max of type I
temp_02 <- temp %>%
  mutate(TypeI = round(TypeI, 2)) %>%
  group_by(type, test, p, size) %>%
  summarise(TypeI = max(TypeI)) %>%
  ungroup() %>%
  arrange(desc(TypeI))

temp_02 %>%
  ggplot(aes(x = p, y = factor(size), fill = TypeI, label = TypeI, group = p)) +
  geom_tile(height = .50, width = .50) +
  geom_text(size = 2.75) +
  scale_x_continuous(breaks = seq(.05, 1, .10), limits = c(0, 1)) +
  scale_fill_gradient(low = "white", high = "red") +
  labs(x = "P", y = "Size", fill = "Type I Error") +
  facet_wrap(vars(type))

```


Both tests performed well in aggregate for all alternative hypothesis.

```{r negativeBonimialTypeI2, echo=FALSE}
temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```

## Geometric
Over the entire range of p, both the exact test and the likelihood ratio  have type I error rate is near zero for some alternative hypothesis. Given the exact test did nearly as bad, the likelihood test succeeded in that is successfully approximated a bad exact test.

```{r geometricTypeI, echo=FALSE}

typeI <- bind_rows(
  readRDS("results/geometric_type_one.rds") %>%
    mutate(type = "Likelihood"),
  readRDS("results/geometric_type_one_exact.rds") %>%
    mutate(type = "Exact")
)

temp <- typeI %>%
  group_by(type, test, alt, p) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(p, test, alt)) %>%
  ungroup()

ggplot(temp, aes(x = p, y = TypeI, group = p)) +
  geom_point() +
  geom_hline(yintercept = .05) +
  # scale_x_continuous(breaks = seq(0.05, 1, .10), limits = c(0, 1)) +
  # scale_y_continuous(breaks = seq(0, 1, .01), limits = c(0, .10)) +
  labs(x = "P", y = "Type I Error") + 
  facet_wrap(vars(test))

```

The test performed very poorly for the greater than alternative. Both tests are extremely conservative.

```{r geometricTypeI2, echo=FALSE}
temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```


