---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Code Overview
Code is split between distribution. Within each distribution (and R file), each parameter has a test. Each test gets a for loop that generates data and runs the test for both type I and type II error. All code ultimately saves a tibble where one row is a one experiment.The gaussian sim file is a good example of how all R files are organized. 

If the code takes a long time to execute, type I and type II error simulations are split. The beta distribution's type I error is split between parameters.

Each script is self contained and should be ran as a background job within RStudio in parallel. Working directory needs to be the folder containing the R project file. 

The sampling distribution folder checks the results produced by the other folders.
