---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```


## Overall Type II Error Rate
Asymptotic type II error rates for MLTesteR's likelihood ratio tests are estimated via simulation. Calculations are based on 5,000 iterations and a sample size of 200. Simulation holds the null hypothesis constant across all alternatives and effect sizes. Where possible, exact tests are included for comparison.

Most tests can achieve near 0% type II error for a large enough effect size. The test for a geometric distribution's p stands out as not getting close to zero. Nearly the entire range of effect sizes was tested for the geometric test and it did not get close to zero.

```{r typeTwoSummary, echo=FALSE, message=FALSE}
library(tidyverse, warn.conflicts= FALSE)
require(scales, warn.conflicts= FALSE)
options(dplyr.summarise.inform = FALSE)

gaussian <- bind_rows(
  readRDS("results/gaussian_type_two.rds"),
  readRDS("results/gaussian_type_two_exact.rds")
)

gamma <- bind_rows(
  readRDS("results/gamma_type_two_rate.rds"),
  readRDS("results/gamma_type_two_scale.rds"),
  readRDS("results/gamma_type_two_shape.rds")
)

poisson <- readRDS("results/poisson_type_two.rds")

beta <- readRDS("results/beta_type_two.rds")

neg_binom <- readRDS("results/negative_binomial_type_two.rds")


geomet <- readRDS("results/geometric_type_two.rds")

expon <- readRDS("results/exponentail_type_two.rds")

binom <- bind_rows(
  readRDS("results/binomail_type_two.rds"),
  readRDS("results/binomial_type_two_exact.rds") %>%
    mutate(test = "binomial_exact")
)

typeII <- bind_rows(
  gaussian,
  gamma, 
  poisson, 
  beta,
  neg_binom,
  geomet,
  expon,
  binom
) %>%
  drop_na()

mark_exact_likelihood <- function(test) {
  type <- case_when(
    str_detect(test, "t.test") ~ "exact",
    str_detect(test, "exact") ~ "exact",
    str_detect(test, "varTest") ~ "exact",
    str_detect(test, "varTest") ~ "exact",
    str_detect(test, "_lr_test") ~ "likelihood",
    TRUE ~ "ERROR")
  
  return(type)
}

typeII <- typeII %>%
  mutate(type = mark_exact_likelihood(test))
# typeII %>% distinct(type, test) %>% arrange(type, test) %>% print(N=Inf)

link_exact_to_likelihood <- function(test) {
  test2 <- case_when(
    str_detect(test, "t.test") ~ "gaussian_mean_lr_test",
    str_detect(test, "varTest") ~ "gaussian_variance_lr_test",
    str_detect(test, "geometric_exact") ~ "geometric_p_lr_test",
    str_detect(test, "negative_binomial_exact") ~ "negative_binomial_p_lr_test",
    str_detect(test, "binomial_exact") ~ "binomial_p_lr_test",
    str_detect(test, "_lr_test") ~ test,
    TRUE ~ "ERROR")
}

typeII <- typeII %>%
  mutate(test2 = link_exact_to_likelihood(test))
# typeII %>% distinct(type, test, test2) %>% arrange(type, test, test2) %>% print(N=Inf)

temp <- typeII %>%
  group_by(type, test2, alt, effectSize) %>%
  summarise(TypeII = mean(pvalue > .05, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeII)) %>%
  ungroup() %>%
  group_by(type, test2) %>%
  summarise(TypeII = min(TypeII)) %>%
  ungroup()

ggplot(temp, aes(x = TypeII, y = test2, color = factor(type))) + 
  geom_point() +
  scale_x_continuous(breaks = seq(0, 1, .10), limits = c(0, 1)) + 
  labs(x = "Minimum Type II Error Rate", y = "Test", color = "Type")

rm(list = ls())


```

# Successful Tests
For a distribution, the likelihood ratio test works well if

* Type II error rates decrease quickly.
* When exact tests are implemented in R, type II error rates are similar to the exact test.

To check the above, one graph is shown per test.


## Gaussian
```{r gaussainTypeII, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
typeII <- bind_rows(
  readRDS("results/gaussian_type_two.rds"),
  readRDS("results/gaussian_type_two_exact.rds")
)

typeII <- typeII %>%
  mutate(test = factor(test, levels = c("gaussian_mean_lr_test", "t.test", "gaussian_variance_lr_test", "varTest")))

temp <- typeII %>%
  group_by(test, alt, effectSize) %>%
  summarise(TypeII = mean(pvalue > .05, na.rm = TRUE)) %>%
  arrange(desc(TypeII)) %>%
  mutate(TypeII = round(TypeII, 2))

temp %>%
  filter(test %in% c("gaussian_mean_lr_test", "t.test")) %>%
  ggplot(aes(x = effectSize, y = TypeII)) +
  geom_point() + 
  scale_x_continuous(breaks = seq(-.50, .50, .20)) +
  scale_y_continuous(breaks = seq(0, 1, .20), limit = c(0, 1)) +
  labs(x = "Effect Size", y = "Type II Error", caption = "Null: mu = 0") + 
  facet_grid(rows = vars(alt), cols = vars(test))

temp %>%
  filter(test %in% c("gaussian_variance_lr_test", "varTest")) %>%
  ggplot(aes(x = effectSize, y = TypeII)) +
  geom_point() + 
  scale_x_continuous(breaks = seq(-5, 5, 2)) +
  scale_y_continuous(breaks = seq(0, 1, .20), limit = c(0, 1)) +
  labs(x = "Effect Size", y = "Type II Error", caption = "Null: variance = 15") + 
  facet_grid(rows = vars(alt), cols = vars(test))

rm(list = ls())

```

## Gamma
```{r gammaTypeII, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
typeII <- bind_rows(
  readRDS(file = "results/gamma_type_two_rate.rds"),
  readRDS(file = "results/gamma_type_two_scale.rds"),
  readRDS(file = "results/gamma_type_two_shape.rds")
)

temp <- typeII %>%
  group_by(test, alt, effectSize) %>%
  summarise(TypeII = mean(pvalue > .05, na.rm = TRUE)) %>%
  arrange(desc(TypeII)) %>%
  mutate(TypeII = round(TypeII, 2))

temp %>%
  ggplot(aes(x = effectSize, y = TypeII)) +
  geom_point() + 
  scale_y_continuous(breaks = seq(0, 1, .20), limit = c(0, 1)) +
  labs(x = "Effect Size", y = "Type II Error") + 
  facet_grid(rows = vars(alt), cols = vars(test), scales = "free_x")

rm(list = ls())

```


## Poisson
```{r poissonTypeII, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
typeII <- readRDS("results/poisson_type_two.rds")

temp <- typeII %>%
  group_by(test, alt, effectSize) %>%
  summarise(TypeII = mean(pvalue > .05, na.rm = TRUE)) %>%
  arrange(desc(TypeII)) %>%
  mutate(TypeII = round(TypeII, 2))

temp %>%
  ggplot(aes(x = effectSize, y = TypeII)) +
  geom_point() + 
  scale_x_continuous(breaks = seq(-.6, .6, .20), labels = comma) +
  scale_y_continuous(breaks = seq(0, 1, .20), limit = c(0, 1)) +
  labs(x = "Effect Size", y = "Type II Error", caption = "Null: lambda = 5") + 
  facet_grid(rows = vars(alt), cols = vars(test))

rm(list = ls())

```

## Beta
```{r betaTypeII, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
typeII <- readRDS("results/beta_type_two.rds")

temp <- typeII %>%
  group_by(test, alt, effectSize) %>%
  summarise(TypeII = mean(pvalue > .05, na.rm = TRUE)) %>%
  arrange(desc(TypeII)) %>%
  mutate(TypeII = round(TypeII, 2))

temp %>%
  ggplot(aes(x = effectSize, y = TypeII)) +
  geom_point() + 
  scale_y_continuous(breaks = seq(0, 1, .20), limit = c(0, 1)) +
  labs(x = "Effect Size", y = "Type II Error") + 
  facet_grid(rows = vars(alt), cols = vars(test))

rm(list = ls())

```

## Exponential
```{r exponentialTypeII, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
typeII <- readRDS("results/exponentail_type_two.rds")

temp <- typeII %>%
  group_by(test, alt, effectSize) %>%
  summarise(TypeII = mean(pvalue > .05, na.rm = TRUE)) %>%
  arrange(desc(TypeII)) %>%
  mutate(TypeII = round(TypeII, 2))

temp %>%
  ggplot(aes(x = effectSize, y = TypeII)) +
  geom_point() + 
  scale_y_continuous(breaks = seq(0, 1, .20), limit = c(0, 1)) +
  labs(x = "Effect Size", y = "Type II Error", caption = "Null: rate = 5") + 
  facet_grid(rows = vars(alt), cols = vars(test))

rm(list = ls())

```

## Binomial
```{r binomTypeII, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
typeII <- bind_rows(
  readRDS("results/binomail_type_two.rds"),
  readRDS("results/binomial_type_two_exact.rds")
)

temp <- typeII %>%
  group_by(test, alt, effectSize) %>%
  summarise(TypeII = mean(pvalue > .05, na.rm = TRUE)) %>%
  arrange(desc(TypeII)) %>%
  mutate(TypeII = round(TypeII, 2),
         effectSize = round(effectSize, 2))

temp %>%
  ggplot(aes(x = effectSize, y = TypeII)) +
  geom_point() + 
  scale_x_continuous(breaks = seq(-.16, .16, .04)) +
  scale_y_continuous(breaks = seq(0, 1, .20), limit = c(0, 1)) +
  labs(x = "Effect Size", y = "Type II Error", caption = "Null: p = .50") + 
  facet_grid(rows = vars(alt), cols = vars(test))

rm(list = ls())

```

## Negative Binomial
```{r negativeBonimialTypeII, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
typeII <- readRDS("results/negative_binomial_type_two.rds")

temp <- typeII %>%
  group_by(test, alt, effectSize) %>%
  summarise(TypeII = mean(pvalue > .05, na.rm = TRUE)) %>%
  arrange(desc(TypeII)) %>%
  mutate(TypeII = round(TypeII, 2),
         effectSize = round(effectSize, 2))

temp %>%
  ggplot(aes(x = effectSize, y = TypeII)) +
  geom_point() + 
  scale_x_continuous(breaks = seq(-.45, .45, .10)) +
  scale_y_continuous(breaks = seq(0, 1, .20), limit = c(0, 1)) +
  labs(x = "Effect Size", y = "Type II Error", caption = "Null: p = .50") + 
  facet_grid(rows = vars(alt), cols = vars(test))

rm(list = ls())
```

# Failures 
For a distribution, the likelihood ratio test is considered bad if

* Type II error rate never hits zero.

## Geometric
For the greater than and two sided alternatives, this test has 100% type II error when the null is .50. It will never reject the null regardless of the data. For all alternatives it never does better than .20%. This test unusable. 

```{r geometricTypeII, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
typeII <- readRDS("results/geometric_type_two.rds")

temp <- typeII %>%
  group_by(test, alt, effectSize) %>%
  summarise(TypeII = mean(pvalue > .05, na.rm = TRUE)) %>%
  arrange(desc(TypeII)) %>%
  mutate(TypeII = round(TypeII, 2),
         effectSize = round(effectSize, 2))

temp %>%
  ggplot(aes(x = effectSize, y = TypeII)) +
  geom_point() + 
  scale_x_continuous(breaks = seq(-.45, .45, .10)) +
  scale_y_continuous(breaks = seq(0, 1, .20), limit = c(0, 1)) +
  labs(x = "Effect Size", y = "Type II Error", caption = "Null: p = .50") + 
  facet_grid(rows = vars(alt), cols = vars(test))

rm(list = ls())
```
