---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

library(tidyverse, warn.conflicts= FALSE)
options(dplyr.summarise.inform = FALSE)

```

# Type I Summary
Asymptotic type I error rates for LRTesteR's likelihood ratio tests are estimated via simulation. Calculations are based on 5,000 iterations and a sample size of 500. Where possible, exact tests are included for comparison.

Most tests have a type I error rate of .05. Exploring each test one by one, most tests have a consistent type I rate across the explored parameter space and alternative hypotheses. The negative binomial has an area where type I error rate increases slightly.

```{r typeOneSummary, echo=FALSE, message=FALSE}

gaussian <- bind_rows(
  readRDS("results/gaussian_type_one.rds"),
  readRDS("results/gaussian_type_one_exact.rds")
)
gaussian <- gaussian %>%
  filter(test %in% c("gaussian_mu_one_sample", "t.test")) %>%
  select(test, alt, param = mu, pvalue) %>%
  bind_rows(gaussian %>%
              filter(test %in% c("gaussian_variance_one_sample", "varTest")) %>%
              select(test, alt, param = variance, pvalue))

gamma <- bind_rows(
  readRDS("results/gamma_type_one_rate.rds"),
  readRDS("results/gamma_type_one_scale.rds"),
  readRDS("results/gamma_type_one_shape.rds")
)
gamma <- gamma %>% 
  filter(test %in% c("gamma_rate_one_sample")) %>%
  select(test, alt, param = rate, pvalue) %>%
  bind_rows(gamma %>% 
              filter(test %in% c("gamma_scale_one_sample")) %>%
              select(test, alt, param = scale, pvalue),
            gamma %>% 
              filter(test %in% c("gamma_shape_one_sample")) %>%
              select(test, alt, param = shape, pvalue))

poisson <- readRDS("results/poisson_type_one.rds")
poisson <- poisson %>%
  select(test, alt, param = lambda, pvalue)


beta <- bind_rows(
  readRDS("results/beta_type_one_shape1.rds"),
  readRDS("results/beta_type_one_shape2.rds")
)
beta <- beta %>% 
  filter(test %in% c("beta_shape1_one_sample")) %>%
  select(test, alt, param = shape1, pvalue) %>%
  bind_rows(beta %>% 
              filter(test %in% c("beta_shape2_one_sample")) %>%
              select(test, alt, param = shape2, pvalue))

neg_binom <- bind_rows(
  readRDS("results/negative_binomial_type_one.rds"),
  readRDS("results/negative_binomial_type_one_exact.rds")  %>%
    mutate(test = "negative_binomial_exact")
)
neg_binom <- neg_binom %>% 
  select(test, alt, param = p, pvalue)

expon <- readRDS("results/exponential_type_one.rds")
expon <- expon %>% 
  select(test, alt, param = rate, pvalue)

binom <- bind_rows(
  readRDS("results/binomail_type_one.rds"),
  readRDS("results/binomial_type_one_exact.rds") %>%
    mutate(test = "binomial_exact")
)
binom <- binom %>% 
  select(test, alt, param = p, pvalue)

cauchy <- readRDS("results/cauchy_type_one.rds")
cauchy <- cauchy %>% 
  filter(test == "cauchy_location_one_sample") %>%
  select(test, alt, param = location, pvalue) %>%
  bind_rows(cauchy %>% 
              filter(test == "cauchy_scale_one_sample") %>%
              select(test, alt, param = scale, pvalue))

invGauus <- readRDS("results/inverse_gaussian_type_one.rds")
invGauus <- invGauus %>% 
  filter(test == "inverse_gaussian_mu_one_sample") %>%
  select(test, alt, param = mu, pvalue) %>%
  bind_rows(invGauus %>% 
              filter(test == "inverse_gaussian_shape_one_sample") %>%
              select(test, alt, param = shape, pvalue),
            invGauus %>% 
              filter(test == "inverse_gaussian_dispersion_one_sample") %>%
              select(test, alt, param = dispersion, pvalue))

empLik <- readRDS("results/empirical_mu_type_one.rds")
empLik <- empLik %>% 
  select(test, alt, param = mu, pvalue)

empLikeQuant <- readRDS("results/empirical_quantile_type_one.rds")
empLikeQuant <- empLikeQuant %>% 
  select(test, alt, param = Q, pvalue)


typeI <- bind_rows(
  gaussian,
  gamma, 
  poisson, 
  beta,
  neg_binom,
  expon,
  binom,
  cauchy,
  invGauus,
  empLik,
  empLikeQuant
)

mark_exact_likelihood <- function(test) {
  type <- case_when(
    str_detect(test, "t.test") ~ "exact",
    str_detect(test, "exact") ~ "exact",
    str_detect(test, "varTest") ~ "exact",
    str_detect(test, "_one_sample") ~ "likelihood",
    TRUE ~ "ERROR")
  
  return(type)
}

typeI <- typeI %>%
  mutate(type = mark_exact_likelihood(test))
# typeI %>% distinct(type, test) %>% arrange(type, test) %>% print(N=Inf)

link_exact_to_likelihood <- function(test) {
  test2 <- case_when(
    str_detect(test, "t.test") ~ "gaussian_mu_one_sample",
    str_detect(test, "varTest") ~ "gaussian_variance_one_sample",
    str_detect(test, "geometric_exact") ~ "geometric_p_one_sample",
    str_detect(test, "negative_binomial_exact") ~ "negative_binomial_p_one_sample",
    str_detect(test, "binomial_exact") ~ "binomial_p_one_sample",
    str_detect(test, "_one_sample") ~ test,
    TRUE ~ "ERROR")
}

typeI <- typeI %>%
  mutate(test2 = link_exact_to_likelihood(test))
# typeI %>% distinct(type, test, test2) %>% arrange(type, test, test2) %>% print(N=Inf)

temp <- typeI %>%
  group_by(type, test2) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

ggplot(temp, aes(x = TypeI, y = test2, color = factor(type))) + 
  geom_point() +
  geom_vline(xintercept = .05) + 
  scale_x_continuous(breaks = seq(-.01, 1.01, .02), limits = c(0, .10)) + 
  labs(x = "Type I Error Rate", y = "Test", color = "Type")

rm(list = ls())
```

# Analysis Criteria
For a distribution, the likelihood ratio test works well if both of the following are true.

* The test has an average of .05 type I error rate over the entire parameter space.
* All tests achieve near .05 type I error for all alternative hypotheses.

To check the above, two graphs are shown per test. For the first point, type I error rate is calculated for many areas in the parameter space.For the second point, results are aggregated across alternative hypotheses.

## Gaussian
```{r gaussainTypeI, echo=FALSE, message=FALSE, results='hide', warning=FALSE}

typeI <- bind_rows(
  readRDS("results/gaussian_type_one.rds"),
  readRDS("results/gaussian_type_one_exact.rds")
)

temp <- typeI %>%
  group_by(test, mu, variance) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup() %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

temp %>%
  ggplot(aes(x = mu, y = variance, fill = TypeI, label = TypeI)) +
  geom_tile(height = .60, width = 1.1) +
  geom_text() +
  scale_x_continuous(breaks = seq(-4, 4, 2)) +
  scale_y_continuous(breaks = seq(1, 5, 2)) +
  scale_fill_gradient(breaks = seq(0, .10, .05), low = "white", high = "red", limit = c(0, .10)) +
  labs(title = "Overall Type I Error Rate", x = "mu", y = "variance", fill = "Type I Error") +
  facet_wrap(vars(test))

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```

## Gamma
```{r gammaTypeI, echo=FALSE}

typeI <- bind_rows(
  readRDS(file = "results/gamma_type_one_rate.rds"),
  readRDS(file = "results/gamma_type_one_scale.rds"),
  readRDS(file = "results/gamma_type_one_shape.rds")
)

temp <- typeI %>%
  group_by(test, shape, rate) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

temp %>%
  ggplot(aes(x = shape, y = rate, fill = TypeI, label = TypeI)) +
  geom_tile(height = .50, width = 1.7) +
  geom_text() +
  scale_x_continuous(breaks = seq(1, 9, 2)) + 
  scale_y_continuous(breaks = seq(1, 9, 2)) + 
  scale_fill_gradient(breaks = seq(0, .10, .05), low = "white", high = "red", limits = c(0, .10)) +
  labs(x = "Shape", y = "Rate", fill = "Type I Error") + 
  facet_wrap(vars(test))

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```


## Poisson
```{r poissonTypeI, echo=FALSE}

typeI <- readRDS("results/poisson_type_one.rds")

temp <- typeI %>%
  group_by(test, lambda) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

ggplot(temp, aes(x = lambda, y = TypeI)) +
  geom_point() +
  geom_hline(yintercept = .05) +
  scale_x_continuous(breaks = seq(0, 100, 1)) +
  scale_y_continuous(breaks = seq(0, 1, .01), limits = c(0, .10)) +
  labs(x = "Lambda", y = "Type I Error")

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```

## Beta
```{r betaTypeI, echo=FALSE}

typeI <- bind_rows(
  readRDS("results/beta_type_one_shape1.rds"),
  readRDS("results/beta_type_one_shape2.rds")
)

temp <- typeI %>%
  group_by(test, shape1, shape2) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

temp %>%
  ggplot(aes(x = shape1, y = shape2, fill = TypeI, label = TypeI)) +
  geom_tile(height = .75, width = 1.1) +
  geom_text() +
  scale_x_continuous(breaks = seq(1, 9, 2)) + 
  scale_y_continuous(breaks = seq(1, 9, 2)) + 
  scale_fill_gradient(breaks = seq(0, .10, .05), low = "white", high = "red", limits = c(0, .10)) +
  labs(x = "Shape 1", y = "Shape 2", fill = "Type I Error") +
  facet_wrap(vars(test))

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```


## Exponential
```{r exponentialTypeI, echo=FALSE}

typeI <- readRDS("results/exponential_type_one.rds")

temp <- typeI %>%
  group_by(test, rate) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

ggplot(temp, aes(x = rate, y = TypeI)) +
  geom_point() +
  geom_hline(yintercept = .05) +
  scale_x_continuous(breaks = seq(0, 100, 1)) +
  scale_y_continuous(breaks = seq(0, 1, .01), limits = c(0, .10)) +
  labs(x = "Rate", y = "Type I Error")

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```

## Binomial
```{r binomTypeI, echo=FALSE}

typeI <- bind_rows(
  readRDS("results/binomail_type_one.rds"),
  readRDS("results/binomial_type_one_exact.rds")
)

temp <- typeI %>%
  group_by(test, p) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(p), test) %>%
  ungroup()

ggplot(temp, aes(x = p, y = TypeI)) +
  geom_point() +
  geom_hline(yintercept = .05) +
  scale_x_continuous(breaks = seq(0.00, 1, .10), limits = c(0, 1)) +
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) +
  labs(x = "P", y = "Type I Error") + 
  facet_wrap(vars(test))

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```


## Negative Binomial
As long as the target number of success is large or p is not near one, the type I error rate is .05. When the target number of successes is small and p is near one, the likelihood test has an increase in type I error rate. How near is too near depends on the target number of successes. Visually this is the bottom right corner of the right graph. 

In the bottom right corner, the exact test is a bit off from the target 5%. It is always conservative while the likelihood test is liberal.

```{r negativeBonimialTypeI, echo=FALSE}

typeI <- bind_rows(
  readRDS("results/negative_binomial_type_one.rds"),
  readRDS("results/negative_binomial_type_one_exact.rds")
)

temp <- typeI %>%
  group_by(test, p, size) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup() %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

temp %>%
  ggplot(aes(x = factor(p), y = factor(size), fill = TypeI, label = TypeI)) +
  geom_tile(height = .60, width = .85) +
  geom_text(size = 2.75) +
  scale_fill_gradient(breaks = seq(0, .10, .05), low = "white", high = "red", limits = c(0, .10)) +
  labs(x = "P", y = "Size", fill = "Type I Error") +
  facet_wrap(vars(test))

```

In the aggregate, the likelihood test performs similarly to other likelihood tests across alternative hypotheses. 

```{r negativeBonimialTypeI2, echo=FALSE}
temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```

## Cauchy
```{r cauchyTypeI, echo=FALSE, message=FALSE, results='hide', warning=FALSE}

typeI <- readRDS("results/cauchy_type_one.rds")

temp <- typeI %>%
  group_by(test, location, scale) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup() %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

temp %>%
  ggplot(aes(x = location, y = scale, fill = TypeI, label = TypeI)) +
  geom_tile(height = .50, width = 1.10) +
  geom_text() +
  scale_x_continuous(breaks = seq(-4, 4, 2)) +
  scale_y_continuous(breaks = seq(1, 9, 2)) +
  scale_fill_gradient(breaks = seq(0, .10, .05), low = "white", high = "red", limit = c(0, .10)) +
  labs(title = "Overall Type I Error Rate", x = "location", y = "scale", fill = "Type I Error") +
  facet_wrap(vars(test))

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```

## Inverse Gaussian
```{r InvGaussI, echo=FALSE}

typeI <- readRDS("results/inverse_gaussian_type_one.rds")

temp <- typeI %>%
  group_by(test, mu, shape) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

temp %>%
  ggplot(aes(x = mu, y = shape, fill = TypeI, label = TypeI)) +
  geom_tile(height = .50, width = 1.65) +
  geom_text() +
  scale_x_continuous(breaks = seq(1, 9, 2)) + 
  scale_y_continuous(breaks = seq(1, 9, 2)) + 
  scale_fill_gradient(breaks = seq(0, .10, .05), low = "white", high = "red", limits = c(0, .10)) +
  labs(x = "Mean", y = "Shape", fill = "Type I Error") + 
  facet_wrap(vars(test))

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```

## Empirical Likelihood For Mu
```{r empTypeI, echo=FALSE}

typeI <- readRDS("results/empirical_mu_type_one.rds")

temp <- typeI %>%
  group_by(test, mu) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

ggplot(temp, aes(x = mu, y = TypeI)) +
  geom_point() +
  geom_hline(yintercept = .05) +
  scale_x_continuous(breaks = seq(-3, 3, 1)) +
  scale_y_continuous(breaks = seq(0, 1, .01), limits = c(0, .10)) +
  labs(x = "Mu", y = "Type I Error")

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```

## Empirical Likelihood For Quantile
```{r empQuantTypeI, echo=FALSE}

typeI <- readRDS("results/empirical_quantile_type_one.rds")

temp <- typeI %>%
  group_by(test, Q) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

ggplot(temp, aes(x = Q, y = TypeI)) +
  geom_point() +
  geom_hline(yintercept = .05) +
  scale_x_continuous(breaks = seq(.05, .95, .10)) +
  scale_y_continuous(breaks = seq(0, 1, .01), limits = c(0, .10)) +
  labs(x = "Quantile", y = "Type I Error")

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```
