---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Type I Summary
Asymptotic type I error rates for MLTesteR's likelihood ratio tests are estimated via simulation. Calculations are based on 5,000 iterations and a sample size of 200. Where possible, exact tests are included for comparison.

Most tests have a type I error rate of .05. Exploring each test one by one, many tests have a consistent type I rate across the entire parameter space and alternative hypotheses. The negative binomial and geometric distributions are exceptions. 

```{r typeOneSummary, echo=FALSE, message=FALSE}
library(tidyverse, warn.conflicts= FALSE)
options(dplyr.summarise.inform = FALSE)

gaussian <- bind_rows(
  readRDS("results/gaussian_type_one.rds"),
  readRDS("results/gaussian_type_one_exact.rds")
)
gaussian <- gaussian %>% 
  filter(test %in% c("gaussian_mean_lr_test", "t.test")) %>%
  select(test, alt, param = mu, pvalue) %>%
  bind_rows(gaussian %>% 
              filter(test %in% c("gaussian_variance_lr_test", "varTest")) %>%
              select(test, alt, param = variance, pvalue))

gamma <- bind_rows(
  readRDS("results/gamma_type_one_rate.rds"),
  readRDS("results/gamma_type_one_scale.rds"),
  readRDS("results/gamma_type_one_shape.rds")
)
gamma <- gamma %>% 
  filter(test %in% c("gamma_rate_lr_test")) %>%
  select(test, alt, param = rate, pvalue) %>%
  bind_rows(gamma %>% 
              filter(test %in% c("gamma_scale_lr_test")) %>%
              select(test, alt, param = scale, pvalue),
            gamma %>% 
              filter(test %in% c("gamma_shape_lr_test")) %>%
              select(test, alt, param = shape, pvalue))

poisson <- readRDS("results/poisson_type_one.rds")
poisson <- poisson %>%
  select(test, alt, param = lambda, pvalue)


beta <- readRDS("results/beta_type_one.rds")
beta <- beta %>% 
  filter(test %in% c("beta_shape1_lr_test")) %>%
  select(test, alt, param = shape1, pvalue) %>%
  bind_rows(beta %>% 
              filter(test %in% c("beta_shape2_lr_test")) %>%
              select(test, alt, param = shape2, pvalue))

neg_binom <- readRDS("results/negative_binomial_type_one.rds")
neg_binom <- neg_binom %>% 
  select(test, alt, param = p, pvalue)

geomet <- readRDS("results/geometric_type_one.rds")
geomet <- geomet %>% 
  select(test, alt, param = p, pvalue)

expon <- readRDS("results/exponentail_type_one.rds")
expon <- expon %>% 
  select(test, alt, param = rate, pvalue)

binom <- bind_rows(
  readRDS("results/binomail_type_one.rds"),
  readRDS("results/binomial_type_one_exact.rds") %>%
    mutate(test = "binomial_exact")
)
binom <- binom %>% 
  select(test, alt, param = p, pvalue)

typeI <- bind_rows(
  gaussian,
  gamma, 
  poisson, 
  beta,
  neg_binom,
  geomet,
  expon,
  binom
) %>%
  drop_na()

mark_exact_likelihood <- function(test) {
  type <- case_when(
    str_detect(test, "t.test") ~ "exact",
    str_detect(test, "exact") ~ "exact",
    str_detect(test, "varTest") ~ "exact",
    str_detect(test, "varTest") ~ "exact",
    str_detect(test, "_lr_test") ~ "likelihood",
    TRUE ~ "ERROR")
  
  return(type)
}

typeI <- typeI %>%
  mutate(type = mark_exact_likelihood(test))
# typeI %>% distinct(type, test) %>% arrange(type, test) %>% print(N=Inf)

link_exact_to_likelihood <- function(test) {
  test2 <- case_when(
    str_detect(test, "t.test") ~ "gaussian_mean_lr_test",
    str_detect(test, "varTest") ~ "gaussian_variance_lr_test",
    str_detect(test, "geometric_exact") ~ "geometric_p_lr_test",
    str_detect(test, "negative_binomial_exact") ~ "negative_binomial_p_lr_test",
    str_detect(test, "binomial_exact") ~ "binomial_p_lr_test",
    str_detect(test, "_lr_test") ~ test,
    TRUE ~ "ERROR")
}

typeI <- typeI %>%
  mutate(test2 = link_exact_to_likelihood(test))
# typeI %>% distinct(type, test, test2) %>% arrange(type, test, test2) %>% print(N=Inf)

temp <- typeI %>%
  group_by(type, test2) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

ggplot(temp, aes(x = TypeI, y = test2, color = factor(type))) + 
  geom_point() +
  geom_vline(xintercept = .05) + 
  scale_x_continuous(breaks = seq(0, 1, .01), limits = c(0, .10)) + 
  labs(x = "Type I Error Rate", y = "Test", color = "Type")

rm(list = ls())
```

# Successful Tests
For a distribution, the likelihood ratio test works well if

* The test has an average of.05 type I error rate over the entire parameter space.
* All tests achieve near .05 type I error for all alternative hypotheses.

To check the above, two graphs are shown per test.

## Gaussian
```{r gaussainTypeI, echo=FALSE, message=FALSE, results='hide', warning=FALSE}

typeI <- bind_rows(
  readRDS("results/gaussian_type_one.rds"),
  readRDS("results/gaussian_type_one_exact.rds")
)

temp <- typeI %>%
  group_by(test, mu, variance) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup() %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

temp %>%
  ggplot(aes(x = mu, y = variance, fill = TypeI, label = TypeI)) +
  geom_tile(height = .50, width = 1) +
  geom_text() +
  scale_x_continuous(breaks = seq(-4, 4, 2)) +
  scale_y_continuous(breaks = seq(1, 5, 2)) +
  scale_fill_gradient(low = "white", high = "red", limit = c(0, .30)) +
  labs(title = "Overall Type I Error Rate", x = "mu", y = "variance", fill = "Type I Error") +
  facet_wrap(vars(test))

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```

## Gamma
```{r gammaTypeI, echo=FALSE}

typeI <- bind_rows(
  readRDS(file = "results/gamma_type_one_rate.rds"),
  readRDS(file = "results/gamma_type_one_scale.rds"),
  readRDS(file = "results/gamma_type_one_shape.rds")
)

temp <- typeI %>%
  group_by(test, shape, rate) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

temp %>%
  ggplot(aes(x = shape, y = rate, fill = TypeI, label = TypeI)) +
  geom_tile(height = .50, width = 1.5) +
  geom_text() +
  scale_x_continuous(breaks = seq(1, 9, 2)) + 
  scale_y_continuous(breaks = seq(1, 9, 2)) + 
  scale_fill_gradient(low = "white", high = "red", limits = c(0, .30)) +
  labs(x = "Shape", y = "Rate", fill = "Type I Error") + 
  facet_wrap(vars(test))

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```


## Poisson
```{r poissonTypeI, echo=FALSE}

typeI <- readRDS("results/poisson_type_one.rds")

temp <- typeI %>%
  group_by(test, lambda) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

ggplot(temp, aes(x = lambda, y = TypeI)) +
  geom_point() +
  geom_hline(yintercept = .05) +
  scale_x_continuous(breaks = seq(0, 100, 1)) +
  scale_y_continuous(breaks = seq(0, 1, .01), limits = c(0, .10)) +
  labs(x = "Lambda", y = "Type I Error")

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```

## Beta
```{r betaTypeI, echo=FALSE}

typeI <- readRDS("results/beta_type_one.rds")

temp <- typeI %>%
  group_by(test, shape1, shape2) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

temp %>%
  ggplot(aes(x = shape1, y = shape2, fill = TypeI, label = TypeI)) +
  geom_tile(height = .75, width = 1.1) +
  geom_text() +
  scale_x_continuous(breaks = seq(1, 9, 2)) + 
  scale_y_continuous(breaks = seq(1, 9, 2)) + 
  scale_fill_gradient(low = "white", high = "red", limits = c(0, .30)) +
  labs(x = "Shape 1", y = "Shape 2", fill = "Type I Error") +
  facet_wrap(vars(test))

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```


## Exponential
```{r exponentialTypeI, echo=FALSE}

typeI <- readRDS("results/exponentail_type_one.rds")

temp <- typeI %>%
  group_by(test, rate) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

ggplot(temp, aes(x = rate, y = TypeI)) +
  geom_point() +
  geom_hline(yintercept = .05) +
  scale_x_continuous(breaks = seq(0, 100, 1)) +
  scale_y_continuous(breaks = seq(0, 1, .01), limits = c(0, .10)) +
  labs(x = "Rate", y = "Type I Error")

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```

## Binomial
```{r binomTypeI, echo=FALSE}

typeI <- bind_rows(
  readRDS("results/binomail_type_one.rds"),
  readRDS("results/binomial_type_one_exact.rds")
)

temp <- typeI %>%
  group_by(test, p) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(p, test)) %>%
  ungroup()

ggplot(temp, aes(x = p, y = TypeI)) +
  geom_point() +
  geom_hline(yintercept = .05) +
  scale_x_continuous(breaks = seq(0.05, 1, .10), limits = c(0, 1)) +
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) +
  labs(x = "P", y = "Type I Error") + 
  facet_wrap(vars(test))

temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```

# Failures 
For a distribution, the likelihood ratio test is considered bad if

* The test does not achieve near .05 type I error rate over any area of the parameter space.
* The test did not achieve near .05 type I error across all alternative hypotheses.

## Negative Binomial
As long as the target number of success is large or p is not near one, the type I error rate is .05. When the target number of successes is small and p is near one, the likelihood test does not have a .05 type I error rate. How near is too near depends on the target number of successes. Visually this is the bottom right corner of the first graph.

```{r negativeBonimialTypeI, echo=FALSE}

typeI <- readRDS("results/negative_binomial_type_one.rds")

temp <- typeI %>%
  group_by(test, p, size) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup() %>%
  mutate(TypeI = round(TypeI, 2),
         meanStat = round(meanStat, 2))

temp %>%
  ggplot(aes(x = factor(p), y = factor(size), fill = TypeI, label = TypeI)) +
  geom_tile(height = .60, width = .80) +
  geom_text(size = 2.75) +
  scale_fill_gradient(low = "white", high = "red", limits = c(0, .30)) +
  labs(x = "P", y = "Size", fill = "Type I Error") +
  facet_wrap(vars(test))

```

In the aggregate, the likelihood test performs similarly to other likelihood tests across alternative hypotheses. 

```{r negativeBonimialTypeI2, echo=FALSE}
temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```

## Geometric
In the above, type I error rates degraded as size decreased. The geometric distribution is the same as the negative binomial with size equal to 1. This suggests the geometric test will do poorly.

Over the entire range of p, the likelihood ratio test has type I error rate far from the desired .05. Sometimes the likelihood test is liberal. Sometimes it is conservative.

```{r geometricTypeI, echo=FALSE}

typeI <- readRDS("results/geometric_type_one.rds")

temp <- typeI %>%
  group_by(test, p) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(p, test)) %>%
  ungroup()

ggplot(temp, aes(x = p, y = TypeI)) +
  geom_point() +
  geom_hline(yintercept = .05) +
  scale_x_continuous(breaks = seq(0.05, 1, .10), limits = c(0, 1)) +
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0,.1)) +
  labs(x = "P", y = "Type I Error") + 
  facet_wrap(vars(test))

```

In the aggregate, type I error rate is below the desired .05 for two alternatives.

```{r geometricTypeI2, echo=FALSE}
temp <- typeI %>%
  group_by(test, alt) %>%
  summarise(TypeI = mean(pvalue <= .05, na.rm = TRUE), meanStat = mean(stat, na.rm = TRUE), N = sum(!is.na(pvalue))) %>%
  arrange(desc(TypeI)) %>%
  ungroup()

temp %>%
  ggplot(aes(x = alt, y = TypeI)) +
  geom_col(alpha = .50) +
  geom_hline(yintercept = .05) + 
  scale_y_continuous(breaks = seq(0, 1, .02), limits = c(0, .10)) + 
  labs(x = "Alternative", y = "Type I Error Rate") +
  facet_wrap(vars(test))

rm(list = ls())

```
